---
title: "Analyses_Parasites_Metapops"
author: "AK Gibson, LT Bubrig"
date: "4/4/2022"
output: html_document
---
Packages
```{r}
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(dplyr)
library(lattice)
library(lme4)
library(MASS)
library(ggplot2)
library(survival)
library(survminer)
library(reshape)
library(logistf)
library(brglm)
```

from glmm_funs.R
```{r}
overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  (rdf <- nrow(model@frame)-model.df)
  rp <- residuals(model)
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE,log.p=TRUE)
  c(chisq=Pearson.chisq,ratio=prat,p=exp(pval),logp=pval)
}
```

Survival analyses

Data import and cleaning
```{r}
# data contains survival for four strains
# for each strain, we had 25 control worms and 25 exposed at start
# tracking survival up to 18 days out

surdat<-read.csv("Bubrig2022_SurvivalData_DRYAD.csv",header=T)
surdat$RandomID=factor(surdat$RandomID) # ID for worm

# removing all censored worms
surdat<-subset(surdat,surdat$Censored!="Y") # remove censored, 7 CB4856, 2 N2, 3 JU2132
sumSur<-subset(surdat,surdat$Death==1) # trim data to day of death

# subset datasets for plotting
sumSurHW<-subset(sumSur,sumSur$Host=="CB4856") # subset
sumSurHW$Host=factor(sumSurHW$Host) # ensure proper leveling

sumSurN2<-subset(sumSur,sumSur$Host=="N2")
sumSurN2$Host=factor(sumSurN2$Host)

sumSur1762<-subset(sumSur,sumSur$Host=="JU1762")
sumSur1762$Host=factor(sumSur1762$Host)

sumSur2132<-subset(sumSur,sumSur$Host=="JU2132")
sumSur2132$Host=factor(sumSur2132$Host)
```

Plotting survival, Fig S2
```{r}
# survival curves
fitHW <- survfit(Surv(Day,Death)~ Treatment, data = sumSurHW) 
fitN2 <- survfit(Surv(Day,Death)~ Treatment, data = sumSurN2)
fit1762 <- survfit(Surv(Day,Death)~ Treatment, data = sumSur1762)
fit2132 <- survfit(Surv(Day,Death)~ Treatment, data = sumSur2132)

a<-ggsurvplot(fitHW, data = sumSurHW,conf.int = TRUE,fun = "pct",ggtheme=theme_minimal(),
           size = 2,  font.tickslab = c(20, "plain", "black"),xlim=c(2,18),break.time.by=4,
           linetype = "strata",palette = c("#0571b0","#ca0020"))
b<-ggsurvplot(fit1762, data = sumSur1762,conf.int = TRUE,fun = "pct",ggtheme=theme_minimal(),
           size = 2,  font.tickslab = c(20, "plain", "black"),xlim=c(2,18),break.time.by=4,
           linetype = "strata",palette = c("#0571b0","#ca0020"))
c<-ggsurvplot(fit2132, data = sumSur2132,conf.int = TRUE,fun = "pct",ggtheme=theme_minimal(),
           size = 2,  font.tickslab = c(20, "plain", "black"),xlim=c(2,18),break.time.by=4,
           linetype = "strata",palette = c("#0571b0","#ca0020"))
d<-ggsurvplot(fitN2, data = sumSurN2,conf.int = TRUE,fun = "pct",ggtheme=theme_minimal(),
           size = 2,  font.tickslab = c(20, "plain", "black"),xlim=c(2,18),break.time.by=4,
           linetype = "strata",palette = c("#0571b0","#ca0020"))
#combine - Fig S2
splots<-list(a,c,b,d)
arrange_ggsurvplots(splots,ncol=2,nrow=2,print=TRUE)
```

Relevant numbers on survival, Table 1
```{r}
# median survival times
med=(matrix(ncol=9,nrow=8))
med<-rbind(summary(fitHW)$table,summary(fit1762)$table,summary(fit2132)$table,summary(fitN2)$table)
rownames(med)<-NULL
host=c(rep("CB4856",2),rep("JU1762",2),rep("JU2132",2),rep("N2",2))
treatment=rep(c("Control","Exposed"),4)
med<-data.frame(cbind(host,treatment,med)) # Table 1

# survival time to days 3,4, and 5 (major reproduction days) - only N2 shows a reduction
summary(survfit(Surv(Day,Death) ~ Treatment, data = sumSurHW), times = 3:5)
summary(survfit(Surv(Day,Death) ~ Treatment, data = sumSur1762), times = 3:5)
summary(survfit(Surv(Day,Death) ~ Treatment, data = sumSur2132), times = 3:5)
summary(survfit(Surv(Day,Death) ~ Treatment, data = sumSurN2), times = 3:5)
```

Analysis for survival - Table S3
```{r}
fit1<-coxph(Surv(Day,Death)~Treatment*Host, data = sumSur) 
fit2<-coxph(Surv(Day,Death)~Treatment+Host, data = sumSur) 
fit3<-coxph(Surv(Day,Death)~Host, data = sumSur)
fit4<-coxph(Surv(Day,Death)~Treatment, data = sumSur)

# model comparison
f.aic<-c(AIC(fit1),AIC(fit2),AIC(fit3),AIC(fit4)) 
delAIC<-f.aic-min(f.aic)
relLik <- exp(-0.5 * delAIC)
aicweight <- relLik/sum(relLik)
aic.table<-data.frame(AIC = f.aic, delAIC = delAIC, relLik = relLik, 
    weight = aicweight)
rownames(aic.table) <- c("host:parasite", "host+parasite", "host", 
    "treatment")
anova(fit1,fit2)

# best model
aic.table
summary(fit1)
```

Fecundity
Data import and cleaning
```{r}
# data contains fecundity data for four strains
# for each, 25 control worms, 25 exposed worms tracked individually
# fecundity counts for 7 days
# for per day fecundity, all data eligible
# for lifetime fecundity, exclude censored worms first
# N2 4 and 48 were also censored (males) but are not in the data

fecdat<-read.csv("Bubrig2022_FecundityData_DRYAD.csv",header=T)
fecdat$RandomizedID=factor(fecdat$RandomizedID) # unique ID within each host genotype
fecdat$UniqueID=factor(fecdat$UniqueID) # each worm has a unique ID
fecdat$ObsID=factor(fecdat$ObsID) # each fecundity count has a unique ID
fecdat$DayID=factor(fecdat$DayID) # day in fecundity collection, unique days by genotype and treatment
fecdat$PlateDay=factor(fecdat$PlateDay)

# exclusion of any censored data points
fecNcen=subset(fecdat,fecdat$Censored!="Y") # remove worms for which data is not available all day, 6 worms total

# sum fecundity across individual worms
SumFec <- fecNcen %>% 
  group_by(UniqueID) %>% 
  summarise( 
    Total = n(),# number of observations
    NOffspring=sum(Count), # total number of offspring per 
    Host=first(Host), # host genotype
    Treatment=first(Treatment)) # exposure treatment

# fecundity schedule - sum by day, host and treatment
FDay<-fecNcen %>% 
  group_by(DayID) %>% # unique for host*treatment to facilitate grouping
  summarise( 
    Total = n(),  
    NOffspring=mean(Count),
    Host=first(Host),
    Treatment=first(Treatment),
    PlateDay=first(PlateDay)) 
total<-tapply(FDay$NOffspring,list(FDay$Host,FDay$Treatment),sum) # mean lifetime fecundity
```

Fecundity Plots, Figure 1 and S3
```{r}
# Figure 1 - lifetime fecundity by genotype and treatment
pf1<-ggplot(SumFec, aes(x=Treatment, y=NOffspring,group=Host,fill=Host,shape=Host)) 
pf1 +
  scale_fill_manual(values=c("#000000","#969696","#d9d9d9","#525252"))+
  stat_summary(fun = mean,position = position_dodge(width = .25),
               geom = "line",size=1.25,linetype="longdash")+
  stat_summary(fun.data = mean_cl_normal,position = position_dodge(width = .25),
               geom = "errorbar",color="black",size=1,width=0.25)+
  stat_summary(fun = mean,position = position_dodge(width = .25),
               geom = "point",cex=7,shape=c(22,22,24,24,23,23,21,21),color="black",
               stroke=1.5)+
  theme_bw()+
  xlab("Treatment")+ylab("Number of Offspring")+
  theme(axis.text.x = element_text(color = "grey20", size = 25),
        axis.text.y = element_text(color = "grey20", size = 25),
        axis.title.y = element_text(color = "grey20", size = 30),
        axis.title.x = element_text(color = "grey20", size = 30))

# Figure S3 - fecundity schedule by genotype and treatment
pfs1<-ggplot(fecdat, aes(x=PlateDay, y=Count,group=Treatment,color=Treatment)) 
pfs1+geom_line(aes(group=UniqueID),alpha=0.2,linetype="solid")+  
  scale_colour_manual(values=c("#0571b0","#ca0020"))+
    facet_grid(cols=vars(Host))+
  theme_bw()+ylim(-1,290)+
    stat_summary(fun = mean,
               geom = "line",size=1.5,aes(group=Treatment,color=Treatment)) +
  stat_summary(fun = mean,
               geom = "linerange",color="black",size=1,
               fun.max = function(x) mean(x) + 
                 qt(.975, df = length(x)) * sd(x) / sqrt(length(x)),
               fun.min = function(x) mean(x) - 
                 qt(.975, df = length(x)) * sd(x) / sqrt(length(x)),
               aes(group=Treatment,color=Treatment))+
   stat_summary(fun = mean,
             geom = "point",cex=4,aes(group=Treatment,color=Treatment))+
  xlab("Day")+ylab("Number of Offspring")+
  theme(axis.text.x = element_text(color = "grey20", size = 25),
        axis.text.y = element_text(color = "grey20", size = 25),
        
        axis.title.y = element_text(color = "grey20", size = 30),
        axis.title.x = element_text(color = "grey20", size = 30))
```

Relevant numbers on fecundity
```{r}
# mean fecundity by treatment and genotype
LFMeans<-tapply(SumFec$NOffspring,list(SumFec$Host,SumFec$Treatment),mean)
LFSE<-tapply(SumFec$NOffspring,list(SumFec$Host,SumFec$Treatment),sd)/sqrt(tapply(SumFec$NOffspring,list(SumFec$Host,SumFec$Treatment),length))

#proportional loss with parasites
propLoss<-1-LFMeans[,2]/LFMeans[,1]
meanpL=mean(propLoss)
sepL=sd(propLoss)/sqrt(4)
#summary of schedule - when are most eggs laid?
FDay$Total=c(rep(LFMeans[1,1],6),rep(LFMeans[1,2],6),rep(LFMeans[4,1],6),rep(LFMeans[4,2],6),
             rep(LFMeans[2,1],6),rep(LFMeans[2,2],6),rep(LFMeans[3,1],6),rep(LFMeans[3,2],6))
dayfraction=FDay$NOffspring/FDay$Total
host<-c(rep("CB4856",12),rep("N2",12),rep("JU1762",12),rep("Ju2132",12))
treatment<-c(rep(c(rep("Control",6),rep("Exposed",6)),2))
day<-c(rep(seq(2,7,1),4))
dailyFraction=data.frame(cbind(host,treatment,day,dayfraction))
colnames(dailyFraction)=c("Host","Treatment", "PlateDay","Proportion of Total")
# see Bubrig2022_DailyFecunditySummaries_DRYAD.xls

# summary of daily fecundity reductions
daily<-1-FDay$NOffspring[FDay$Treatment=="Exposed"]/FDay$NOffspring[FDay$Treatment=="Control"]
host<-c(rep("CB4856",6),rep("N2",6),rep("JU1762",6),rep("Ju2132",6))
dailyLoss<-data.frame(cbind(host,day,daily))
colnames(dailyLoss)=c("Host","PlateDay","Proportional Reduction")

```

Analysis for fecundity, Tables S4,S5
```{r}
#Lifetime fecundity
# fixed = host, treatment
# distribution = poisson
lfmodel1<-glm(NOffspring~Host*Treatment,data=SumFec,family="poisson")
dp = sum(residuals(lfmodel1,type ="pearson")^2)/lfmodel1$df.residual
dp # heavily overdispersed
# negative binomial
lfmodel3<-glm.nb(NOffspring~Host*Treatment,data=SumFec)
AIC(lfmodel1,lfmodel3)

# other candidate models
lfmodel4=update(lfmodel3,.~.-Host:Treatment) # evaluate interaction
lfmodel5=update(lfmodel4,.~.-Treatment) # treatment
lfmodel6=update(lfmodel4,.~.-Host) # host

# model comparison
f.aic<-c(lfmodel3$aic,lfmodel4$aic,lfmodel5$aic,lfmodel6$aic) 
delAIC<-f.aic-min(f.aic)
relLik <- exp(-0.5 * delAIC)
aicweight <- relLik/sum(relLik)
aic.table<-data.frame(AIC = f.aic, delAIC = delAIC, relLik = relLik, 
    weight = aicweight)
rownames(aic.table) <- c("host:parasite", "host+parasite", "host", 
    "treatment")
anova(lfmodel3,lfmodel4)

# model 3 with the interaction, indicating differential reactions across genotypes
aic.table
summary(lfmodel3) 

# fecundity schedule
# fixed = host * treatment, treatment*Day
# random = Unique ID
# distribution = poisson
fs1<-glmer(Count~Treatment*Host+ Treatment*PlateDay+(1|UniqueID),
           data=fecdat,family="poisson",control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
overdisp_fun(fs1)
fs2<-glmer(Count~Treatment*Host + Treatment*PlateDay+(1|UniqueID) + (1|ObsID),
           data=fecdat,family="poisson",control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
fs3<-update(fs2,.~.-Treatment:Host)
fs4<-update(fs2,.~.-Treatment:PlateDay)
fs5<-update(fs3,.~.-Treatment:PlateDay)

f.aic<-c(AIC(fs2),AIC(fs3),AIC(fs4),AIC(fs5)) 
delAIC<-f.aic-min(f.aic)
relLik <- exp(-0.5 * delAIC)
aicweight <- relLik/sum(relLik)
aic.table<-data.frame(AIC = f.aic, delAIC = delAIC, relLik = relLik, 
    weight = aicweight)
rownames(aic.table) <- c("treatment*Host+treatment*day","treatment*day","treatment*host",
                         "Treatment+Host+Day")
aic.table
summary(fs2)
```

Population growth
Data import and cleaning
```{r}
# dataset contains 3 blocks
# Block 2 - N2 and CB4856
# Block 3 - 5 "wild" lines 
# Block 4 - N2, CB4856, + 4 "wild" lines
# Worm #76 in block 4, JU2816 control, censored because of errors in processing plates - not in the dataset

pgdat<-read.csv("Bubrig2022_PopulationGrowth_DRYAD.csv",header=T)
pgdat<-subset(pgdat,pgdat$Male!="Y") # excluding populations founded by males
pgdat<-transform(pgdat,
                 Randomized.ID=factor(Randomized.ID), # ID within block
                 Block=factor(Block), # separate trials
                 Count.Replicate=factor(Count.Replicate), # replicate counts within worms
                 ObsID=factor(ObsID),
                 Host=factor(Host)) # id for each individual count

# add an id term
pgdat<-within(pgdat,
              {
                ID<-interaction(Block, Randomized.ID)
              }) # unique ID for each population, across all blocks

# mean for populations
pg_sum1 <- pgdat %>% 
  group_by(ID) %>% 
  summarise( 
    Total = n(),  
    mean.count=mean(Count),
    Host=first(Host),
    Treatment=first(Treatment)) 

# summary
PGMeans<-tapply(pg_sum1$mean.count,list(pg_sum1$Host,pg_sum1$Treatment),mean,na.rm=T)
PGSD<-tapply(pg_sum1$mean.count,list(pg_sum1$Host,pg_sum1$Treatment),sd,na.rm=T)
PGSE<-PGSD/sqrt(tapply(pg_sum1$mean.count,list(pg_sum1$Host,pg_sum1$Treatment),length))
ratio = PGMeans[,2]/PGMeans[,1]
ratioSE = ratio*sqrt((PGSE[,1]/PGMeans[,1])^2+(PGSE[,2]/PGMeans[,2])^2)
f2dat=cbind(ratio,ratioSE)
f2dat <- cbind(rownames(f2dat), data.frame(f2dat, row.names=NULL))
colnames(f2dat)=c("Host","Ratio","SE")
```

Population Growth Plot - Figure 2, S4
```{r}
#Figure 2
pf2<-ggplot(pg_sum1, aes(x=Treatment, y=mean.count,group=Host,fill=Host,shape=Host)) 
pf2 +
  scale_fill_manual(values=c("#000000","white","#969696","#d9d9d9","white","white","#525252"))+
  stat_summary(fun = mean,position = position_dodge(width = .25),
               geom = "line",size=c(1.25,1.25,0.75,0.75,1.25,1.25,1.25,1.25,0.75,0.75,0.75,0.75,1.25,1.25),linetype=c("solid","solid","dotted","dotted","solid","solid","solid","solid","dotted","dotted","dotted","dotted","solid","solid"))+
  stat_summary(fun.data = mean_cl_normal,position = position_dodge(width = .25),
               geom = "errorbar",color="black",size=1,width=0.25)+
  stat_summary(fun = mean,position = position_dodge(width =0.25),
               geom = "point",cex=c(7,7,4,4,7,7,7,7,4,4,4,4,7,7),shape=c(22,22,25,25,24,24,23,23,22,22,21,21,21,21),color="black",
               stroke=1.5)+
  theme_bw()+
  xlab("Treatment")+ylab("Number of Offspring")+
  theme(axis.text.x = element_text(color = "grey20", size = 25),
        axis.text.y = element_text(color = "grey20", size = 25),
        axis.title.y = element_text(color = "grey20", size = 30),
        axis.title.x = element_text(color = "grey20", size = 30))

#Figure S4
#showing data for each genotype
# order
pg_sum1o=pg_sum1
pg_sum1o$Host=factor(pg_sum1$Host,levels=c("CB4856","JU1762","JU2132","N2","JU1249","JU2287","JU2816"))
pfs2<-ggplot(pg_sum1o, aes(x=Treatment, y=mean.count,group=Treatment,color=Treatment)) 
pfs2+ geom_jitter(position=position_jitter(0.2),cex=4,alpha=0.6) +
  scale_colour_manual(values=c("#0571b0","#ca0020"))+
  stat_summary(fun = mean,
               geom = "linerange",color="black",size=1,
               fun.max = function(x) mean(x) + 
                 qt(.975, df = length(x)) * sd(x) / sqrt(length(x)),
               fun.min = function(x) mean(x) - 
                 qt(.975, df = length(x)) * sd(x) / sqrt(length(x)))+
  stat_summary(fun = mean,
               geom = "point",cex=5,color="black")+
  stat_summary(fun = mean,
               geom = "line",size=1,color="black",aes(group=Host))+
  facet_grid(cols=vars(Host))+
  theme_bw()+
  theme(axis.text.x = element_text(color = "grey20", size = 15),
        axis.text.y = element_text(color = "grey20", size = 15),
        axis.title.y = element_text(color = "grey20", size = 20),
        axis.title.x = element_text(color = "grey20", size = 20))
```

Population Growth Relevant Numbers
```{r}
PGTable<-cbind(PGMeans[,1],PGSE[,1],PGMeans[,2],PGSE[,2],ratio,ratioSE)
colnames(PGTable)=c("Control Mean","Control SE","Exposed Mean","Exposed SE","Ratio","RatioSE")
1-mean(PGTable[,5])
sd(PGTable[,5])/sqrt(7)
1-PGTable[,5]
```

Analysis of Population Growth, Table S6
```{r}
# fixed effect: host, treatment, and interaction
# random effects: block and unique ID to group counts from same individual
pgmodel1<-glmer(Count~Host*Treatment+Block+(1|ID),data=pgdat,family="poisson")
overdisp_fun(pgmodel1)  # overdispersed
pgmodel2 <- glmer(Count~Host*Treatment+Block+(1|ID)+(1|ObsID),data=pgdat,family="poisson",
                  control=glmerControl(optimizer="bobyqa",
                            optCtrl=list(maxfun=2e5)))# add in observation id to deal with overdispersion
overdisp_fun(pgmodel2)

pgmodel3<-update(pgmodel2,.~.-Host:Treatment)
pgmodel4<-update(pgmodel3,.~.-Host)
pgmodel5<-update(pgmodel3,.~.-Treatment)

# model comparison
f.aic<-c(AIC(pgmodel2),AIC(pgmodel3),AIC(pgmodel4),AIC(pgmodel5)) 
delAIC<-f.aic-min(f.aic)
relLik <- exp(-0.5 * delAIC)
aicweight <- relLik/sum(relLik)
aic.table<-data.frame(AIC = f.aic, delAIC = delAIC, relLik = relLik, 
    weight = aicweight)
rownames(aic.table) <- c("host:parasite", "host+parasite", "treatment", 
    "host")

# model 2 with the interaction, indicating differential reactions across genotypes
aic.table
summary(pgmodel2) 
summary(pgmodel3)  
```

Dispersal
Data import and cleaning
```{r}
# Experiment 1
  # data contains dispersal success data for N2
  # control: 2 populations, 20 worms each
  # exposed: 3 populations, 20 worms each

OD<-read.csv("Bubrig2022_DispersalExp1_DRYAD.csv",header=T)
OD<-subset(OD,OD$Censored==0)

# summarize
# count number of worms that made babies for each treatment*dose*replicate
SumOD <- OD %>% 
  group_by(PopID) %>% # ID grouping individuals with same trial*dose
  summarise( 
    Total = n(),  
    Babies=sum(Babies),
    Treatment=first(Treatment),
    Dose=first(EffectiveDose))
SumOD$Rate = SumOD$Babies/SumOD$Total

# Dispersal Experiment 2
  # 3 doses/timings of exposure
  # 1 genotype, N2
  # 3 populations each, 20 worms per population
DD<-read.csv("Bubrig2022_DispersalExp2_DRYAD.csv",header=T)
DD$PlateID=factor(DD$PlateID)
DD$ExposureDay=factor(DD$ExposureDay) # dose - order may be confusing
DD$Replicate=factor(DD$Replicate)

# exclude
# there is one male in this dataset, in the control treatment 
# 1 exposed and 1 control worm were censored
DD<-subset(DD,DD$Male!=1 & DD$Censored!=1)

# summarize
DD = subset(DD,DD$Babies!="NA") # removes data points before decision point
# count number of worms that made babies for each treatment*dose*replicate
SumDD <- DD %>% 
  group_by(RepID) %>% # identifies worms from same parent replicate plate
  summarise( 
    Total = n(), # total worms
    Babies=sum(Babies)) # total worms that made babies

# adding additional information to table - dose, treatment, success rate
Dose=rep(c("high","high","high","medium","medium","medium","low","low","low"),2)
Treatment=c(rep("Control",9),rep("Exposed",9))
SuccessRate = SumDD$Babies/SumDD$Total
SumDD<-cbind(SumDD,Dose,Treatment,SuccessRate)

# Experiment 3
  # 1 dose/timing of exposure - medium
  # 4 genotypes, N2, CB4856, JU1762, JU2132
  # 3 populations each, 20 worms each
DV<-read.csv("Bubrig2022_DispersalExp3_DRYAD.csv",header=T)
DV$PlateID=factor(DV$PlateID)
DV$Replicate=factor(DV$Replicate)
DV$RepID=factor(DV$RepID)

# exclude censored individuals
DV<-subset(DV,DV$Censored!=1)
# DV includes males
# DVf does not include males
DVf<-subset(DV,DV$Male!=1)

# subset on data points that establish reproductive success
DV1 = subset(DV, DV$BabiesAlt!="NA") # with males, coded as successful
DVf1 = subset(DVf,DVf$Babies!="NA") # without males

SumDV <- DV1 %>% 
  group_by(RepID) %>%
  summarise( 
    Total = n(),  
    Reproductive=sum(BabiesAlt),
    Host=first(Host),
    Treatment=first(Treatment)) 
SumDV$SuccessRate=SumDV$Reproductive/SumDV$Total
SumDVf <- DVf1 %>% 
  group_by(RepID) %>%
  summarise( 
    Total = n(),  
    Reproductive=sum(Babies),
    Host=first(Host),
    Treatment=first(Treatment)) 
SumDVf$SuccessRate=SumDVf$Reproductive/SumDVf$Total

```

Dispersal Success Plots, Figure 3, S5
```{r}
# Figure 3A - experiment 1, dispersal success by treatment
pf3a<-ggplot(SumOD, aes(x=Treatment, y=Rate,group=Treatment)) 
pf3a +
  stat_summary(fun.data = mean_se,
               geom = "errorbar",color="black",size=1,width=0.35)+ylim(-0.025,1.025)+
  stat_summary(fun = base::mean,geom = "point",cex=9,shape=21,stroke=1.5, fill="#525252")+
  theme_bw()+
  xlab("Treatment")+ylab("Rate")+
  theme(axis.text.x = element_text(color = "grey20", size = 20),
        axis.text.y = element_text(color = "grey20", size = 25),
        axis.title.y = element_text(color = "grey20", size = 30),
        axis.title.x = element_text(color = "grey20", size = 30))

# Figure 3b - experiment 2, dispersal success by exposure level
SumDD$Dose=factor(SumDD$Dose,levels=c("high","medium","low"))
pf3b<-ggplot(SumDD, aes(x=Treatment,y=SuccessRate,group=Treatment,fill=Treatment)) 
pf3b + ylim(-0.025,1.025)+
  stat_summary(fun.data = mean_se,
               geom = "errorbar",color="black",size=1,width=0.35)+
   stat_summary(fun = base::mean,
               geom = "point",cex=9,shape=21,stroke=1.5, fill="#525252")+
  theme_bw()+ facet_grid(cols=vars(Dose))+
  xlab("Treatment")+ylab("Rate")+
  theme(axis.text.x = element_text(color = "grey20", size = 20),
        axis.text.y = element_text(color = "grey20", size = 25),
        axis.title.y = element_text(color = "grey20", size = 30),
        axis.title.x = element_text(color = "grey20", size = 30))

# Figure 3c - experiment 3, dispersal success by treatment and genotype, without males
pf3c<-ggplot(SumDVf, aes(x=Treatment,y=SuccessRate,group=Treatment,fill=Host)) 
pf3c + ylim(-0.025,1.025)+
  scale_fill_manual(values=c("#000000","#969696","#d9d9d9","#525252"))+
  stat_summary(fun.data = mean_se,
               geom = "errorbar",color="black",size=1,width=0.35)+
   stat_summary(fun = base::mean,
               geom = "point",cex=9,shape=c(22,22,24,24,23,23,21,21),stroke=1.5)+
  theme_bw()+ facet_grid(cols=vars(Host))+
  xlab("Treatment")+ylab("Rate")+
  theme(axis.text.x = element_text(color = "grey20", size = 20),
        axis.text.y = element_text(color = "grey20", size = 25),
        axis.title.y = element_text(color = "grey20", size = 30),
        axis.title.x = element_text(color = "grey20", size = 30))

# Figure S5 - figure 3C with males
pfs3c<-ggplot(SumDV, aes(x=Treatment,y=SuccessRate,group=Treatment,fill=Host)) 
pfs3c + ylim(-0.025,1.025)+
  scale_fill_manual(values=c("#000000","#969696","#d9d9d9","#525252"))+
  stat_summary(fun.data = mean_se,
               geom = "errorbar",color="black",size=1,width=0.35)+
   stat_summary(fun = base::mean,
               geom = "point",cex=9,shape=c(22,22,24,24,23,23,21,21),stroke=1.5)+
  theme_bw()+ facet_grid(cols=vars(Host))+
  xlab("Treatment")+ylab("Rate")+
  theme(axis.text.x = element_text(color = "grey20", size = 20),
        axis.text.y = element_text(color = "grey20", size = 25),
        axis.title.y = element_text(color = "grey20", size = 30),
        axis.title.x = element_text(color = "grey20", size = 30))
```

Dispersal Relevant Numbers
```{r}
# Experiment 1
total<-melt(tapply(SumOD$Total,list(SumOD$Treatment),sum))
babies<-melt(tapply(SumOD$Babies,list(SumOD$Treatment),sum))
mean<-melt(tapply(SumOD$Rate,list(SumOD$Treatment),mean))
se<-melt(tapply(SumOD$Rate,list(SumOD$Treatment),sd)/sqrt(tapply(SumOD$Rate,list(SumOD$Treatment),length)))
ODSum = cbind(mean,se$value,total$value,babies$value)
colnames(ODSum) = c("Treatment","Mean","SE","Total","TotalBabies")

# Experiment 2 - SumDD for counts; DDSum for mean success rates
# means and errors
mean<-melt(tapply(SumDD$SuccessRate,list(SumDD$Dose,SumDD$Treatment),mean))
se<-melt(tapply(SumDD$SuccessRate,list(SumDD$Dose,SumDD$Treatment),sd)/sqrt(3))
DDSum = cbind(mean,se$value)
colnames(DDSum) = c("Dose","Treatment","Mean","SE")
DDSum$Dose=factor(DDSum$Dose,levels=c("high","medium","low"))
mean(mean$value[mean$X2=="Exposed"])
sd(mean$value[mean$X2=="Exposed"])/sqrt(3)

# Experiment 3
# means and errors
# without males
meanDVf<-melt(tapply(SumDVf$SuccessRate,list(SumDVf$Host,SumDVf$Treatment),mean))
seDVf<-melt(tapply(SumDVf$SuccessRate,list(SumDVf$Host,SumDVf$Treatment),sd)/sqrt(3))
DVfSum = cbind(meanDVf,seDVf$value)
colnames(DVfSum) = c("Host","Treatment","Mean","SE")
mean(DVfSum$Mean[DVfSum$Treatment=="Exposed"])
sd(DVfSum$Mean[DVfSum$Treatment=="Exposed"])/sqrt(4)
# with males
meanDV<-melt(tapply(SumDV$SuccessRate,list(SumDV$Host,SumDV$Treatment),mean))
seDV<-melt(tapply(SumDV$SuccessRate,list(SumDV$Host,SumDV$Treatment),sd)/sqrt(3))
DVSum = cbind(meanDV,seDV$value)
colnames(DVSum) = c("Host","Treatment","Mean","SE")
```

Analysis - tables s7-10
```{r}
# Experiment 1 Table S7
# models
om1<-brglm(cbind(Babies,Total-Babies)~Treatment,data=SumOD,family=binomial(logit),method="glm.fit") # issue with rare events
om2<-brglm(cbind(Babies,Total-Babies)~Treatment,data=SumOD,family=binomial(logit),method="brglm.fit") 
# om2 has more reasonable estimates
om3<-update(om2,.~.-Treatment)
# model comparison
f.aic<-c(AIC(om2),AIC(om3)) 
delAIC<-f.aic-min(f.aic)
relLik <- exp(-0.5 * delAIC)
aicweight <- relLik/sum(relLik)
aic.table<-data.frame(AIC = f.aic, delAIC = delAIC, relLik = relLik, 
    weight = aicweight)
rownames(aic.table) <- c("+treatment","-treatment")
# winner - Table S7
aic.table
summary(om2)

# Experiment 2 Table S8
# models
dd1<-brglm(cbind(Babies,Total-Babies)~Dose*Treatment,data=SumDD,family=binomial(logit),method="glm.fit") 
dd2<-brglm(cbind(Babies,Total-Babies)~Dose*Treatment,data=SumDD,family=binomial(logit),method="brglm.fit")
dd3<-update(dd2,.~.-Dose:Treatment)
dd4<-update(dd3,.~.-Treatment)
dd5<-update(dd3,.~.-Dose)
dd6<-update(dd5,.~.-Treatment)
# model comparison
f.aic<-c(AIC(dd2),AIC(dd3),AIC(dd4),AIC(dd5) )
delAIC<-f.aic-min(f.aic)
relLik <- exp(-0.5 * delAIC)
aicweight <- relLik/sum(relLik)
aic.table<-data.frame(AIC = f.aic, delAIC = delAIC, relLik = relLik, 
    weight = aicweight)
rownames(aic.table) <- c("Dose:Treatment","Dose+Treatment","Dose","Treatment")
# winner
aic.table
summary(dd3)

# Experiment 3 - Table S9
# without males
dv1<-brglm(cbind(Reproductive,Total-Reproductive)~Host*Treatment,data=SumDVf,family=binomial(logit),method="glm.fit") 
# bias-reduced fit
dv2<-brglm(cbind(Reproductive,Total-Reproductive)~Host*Treatment,data=SumDVf,family=binomial(logit),method="brglm.fit")
# parameters look better
dv3<-update(dv2,.~.-Host:Treatment)
dv4<-update(dv3,.~.-Treatment)
dv5<-update(dv3,.~.-Host)
# model comparison
f.aic<-c(AIC(dv2),AIC(dv3),AIC(dv4),AIC(dv5)) 
delAIC<-f.aic-min(f.aic)
relLik <- exp(-0.5 * delAIC)
aicweight <- relLik/sum(relLik)
aic.table<-data.frame(AIC = f.aic, delAIC = delAIC, relLik = relLik, 
    weight = aicweight)
rownames(aic.table) <- c("Host:Treatment","Host+Treatment","Host","Treatment")
aic.table
summary(dv3)
summary(dv2)

# with males - table S10
dvm1<-brglm(cbind(Reproductive,Total-Reproductive)~Host*Treatment,data=SumDV,family=binomial(logit),method="glm.fit") 
# bias-reduced fit
dvm2<-brglm(cbind(Reproductive,Total-Reproductive)~Host*Treatment,data=SumDV,family=binomial(logit),method="brglm.fit")
# parameters look better
dvm3<-update(dvm2,.~.-Host:Treatment)
dvm4<-update(dvm3,.~.-Treatment)
dvm5<-update(dvm3,.~.-Host)
# model comparison
f.aic<-c(AIC(dvm2),AIC(dvm3),AIC(dvm4),AIC(dvm5)) 
delAIC<-f.aic-min(f.aic)
relLik <- exp(-0.5 * delAIC)
aicweight <- relLik/sum(relLik)
aic.table<-data.frame(AIC = f.aic, delAIC = delAIC, relLik = relLik, 
    weight = aicweight)
rownames(aic.table) <- c("Host:Treatment","Host+Treatment","Host","Treatment")
aic.table
summary(dvm3)
summary(dvm2)
```

Project of parasite-mediated selection across scales
Data
```{r}
# Fecundity - fecundity of exposed relative to control hermaphrodites
FRatio<-LFMeans[,2]/LFMeans[,1]
FRatioSE<-FRatio*sqrt((LFSE[,1]/LFMeans[,1])^2+(LFSE[,2]/LFMeans[,2])^2)

# Population Growth - population size of exposed relative to control hermaphrodites
PGRatio = PGMeans[,2]/PGMeans[,1]
PGRatioSE = PGRatio*sqrt((PGSE[,1]/PGMeans[,1])^2+(PGSE[,2]/PGMeans[,2])^2)
# reduce to focal relevant genotypes
PGRatio=c(PGRatio[1],PGRatio[3],PGRatio[4],PGRatio[7])
PGRatioSE=c(PGRatioSE[1],PGRatioSE[3],PGRatioSE[4],PGRatioSE[7])

# Dispersal - relative number of successful colonists of exposed relative to control - no males
DVfM=cbind(DVfSum$Mean[1:4],DVfSum$Mean[5:8])
rownames(DVfM)=DVfSum$Host[1:4]
colnames(DVfM)=c("Control","Exposed")
DVfSE=cbind(DVfSum$SE[1:4],DVfSum$SE[5:8])
rownames(DVfSE)=DVfSum$Host[1:4]
colnames(DVfSE)=c("Control","Exposed")
PGM=PGMeans[-c(2,5,6),]
PGSE2=PGSE[-c(2,5,6),]

Colonist = PGM*DVfM
ColonistSE = Colonist * sqrt((PGSE2 / PGM) ^ 2 + (DVfSE / DVfM) ^ 2)

DfRatio<-Colonist[,2]/Colonist[,1]
DfRatioSE<-DfRatio*sqrt((ColonistSE[,1]/Colonist[,1])^2+(ColonistSE[,2]/Colonist[,2])^2)

# Dispersal - relative number of successful colonists of exposed relative to control - with males
DVM=cbind(DVSum$Mean[1:4],DVSum$Mean[5:8])
rownames(DVM)=DVSum$Host[1:4]
colnames(DVM)=c("Control","Exposed")
DVSE=cbind(DVSum$SE[1:4],DVSum$SE[5:8])
rownames(DVSE)=DVSum$Host[1:4]
colnames(DVSE)=c("Control","Exposed")

Colonistm = PGM*DVM
ColonistmSE = Colonistm * sqrt((PGSE2 / PGM) ^ 2 + (DVSE / DVM) ^ 2)

DRatio<-Colonistm[,2]/Colonistm[,1]
DRatioSE<-DRatio*sqrt((ColonistmSE[,1]/Colonistm[,1])^2+(ColonistmSE[,2]/Colonistm[,2])^2)

# making data frames
Host=c("JU2132","JU1762","N2","CB4856")
index<-c(4,2,1,3)
#Fecundity
FMat=data.frame(rep("Fecundity",4),FRatio[order(index)],FRatioSE[order(index)])
FMat<-cbind(Host,FMat,row.names=NULL)
colnames(FMat)=c("Host","Metric","Ratio","SE")
#PopulationGrowth
PGMat=data.frame(rep("Population Growth",4),PGRatio[order(index)],PGRatioSE[order(index)])
PGMat<-cbind(Host,PGMat,row.names=NULL)
colnames(PGMat)=c("Host","Metric","Ratio","SE")
#Dispersal
#no males
DfMat=data.frame(rep("Dispersal",4),DfRatio[order(index)],DfRatioSE[order(index)])
DfMat<-cbind(Host,DfMat,row.names=NULL)
colnames(DfMat)=c("Host","Metric","Ratio","SE")
#with males
DMat=data.frame(rep("Dispersal",4),DRatio[order(index)],DRatioSE[order(index)])
DMat<-cbind(Host,DMat,row.names=NULL)
colnames(DMat)=c("Host","Metric","Ratio","SE")

DComF=rbind(FMat,PGMat,DfMat)
DCom=rbind(FMat,PGMat,DMat)
DComF$Metric=factor(DComF$Metric,levels=c("Fecundity","Population Growth","Dispersal"))
DComF$Host=factor(DComF$Host,levels=c("CB4856","N2","JU1762","JU2132"))
DCom$Metric=factor(DCom$Metric,levels=c("Fecundity","Population Growth","Dispersal"))
DCom$Host=factor(DCom$Host,levels=c("CB4856","N2","JU1762","JU2132"))

# relevant comparisons
fr<-mean(DComF$Ratio[DComF$Metric=="Fecundity"])
sd(DComF$Ratio[DComF$Metric=="Fecundity"])/sqrt(4)
pgr<-mean(DComF$Ratio[DComF$Metric=="Population Growth"])
sd(DComF$Ratio[DComF$Metric=="Population Growth"])/sqrt(4)
dr<-mean(DComF$Ratio[DComF$Metric=="Dispersal"])
sd(DComF$Ratio[DComF$Metric=="Dispersal"])/sqrt(4)

(1-dr)/(1-fr) # dispersal performance reduction against fecundity
(1-dr)/(1-pgr) # dispersal performance reduction against population growth
#with males
dr2<-mean(DCom$Ratio[DCom$Metric=="Dispersal"])
(1-dr2)/(1-fr)
(1-dr2)/(1-pgr)

#comparison of other genotypes to cb4856, "resistant" genotype
fsr=mean(DComF[5:7,3]) #CB4856 DCOMf ratio for popG
0.72065670/fsr
fsd=mean(DComF[9:11,3]) #CB4856 DCOMf ratio for fecundity
0.27672275/fsd # DComF, dispersal CB4856/mean of other genotypes
#with males
fsdM=mean(DCom[9:11,3])
0.46126243/fsd
```

Figures of cross-scale comparison - Figure 4, s6
```{r}
palette=c("#000000","#525252","#969696","#d9d9d9")

#without males - Fig. 4
p4<-ggplot(data=DComF,aes(y=Host,x=Ratio,fill=Host))
p4+facet_grid(cols=vars(Metric))+
  geom_errorbar(aes(xmin=Ratio-SE, xmax=Ratio+SE), width=.2,size=1.2,color="black")+
  geom_bar(stat="identity")+
  scale_fill_manual(values=palette)+
  xlim(0,1)+
  theme_minimal()+
  theme(axis.text.x = element_text(color = "grey20", size = 20),
        axis.text.y = element_text(color = "grey20", size = 20),
        axis.title.y = element_text(color = "grey20", size = 25),
        axis.title.x = element_text(color = "grey20", size = 25))+
  theme(panel.spacing = unit(2, "lines"))


#without males - Fig S6
pS4<-ggplot(data=DCom,aes(y=Host,x=Ratio,fill=Host))
pS4+facet_grid(cols=vars(Metric))+
  geom_errorbar(aes(xmin=Ratio-SE, xmax=Ratio+SE), width=.2,size=1.2,color="black")+
  geom_bar(stat="identity")+
  scale_fill_manual(values=palette)+
  xlim(0,1)+
  theme_minimal()+
  theme(axis.text.x = element_text(color = "grey20", size = 20),
        axis.text.y = element_text(color = "grey20", size = 20),
        axis.title.y = element_text(color = "grey20", size = 25),
        axis.title.x = element_text(color = "grey20", size = 25))+
  theme(panel.spacing = unit(2, "lines"))
```

